---
title: "Bikeshare data Project"
author: "Adeyinka Abigail Aremo"
date: "2/8/2020"
output:
pdf_document: default
html_document:
  df_print: paged
---

```{r setup, include=FALSE}
```


```{r RUNNING PACKAGES AND READING FILES}
library(ggplot2)                     #Calling the ggplot Library
library(dplyr)                       #Calling the dplyr Library
library(lubridate)                   #Calling the lubridate Library
library(tidyr)                       #Calling the tidyr Library

ny = read.csv('new-york-city.csv')
head(ny)

wash = read.csv('washington.csv')
head(wash)

chi = read.csv('chicago.csv')
head(chi)
```

#NUMBER 1.TRIP DURATION
# Write a function to determine the mean of each dataset trip duration and determine the User types
#that accumulately the most trip duration? 

```{r MEAN FUNCTION FOR CITIES}
#1a. Mean Function
#Writing the Mean Function
mean.function=function(X){      #Writing a mean function
M=sum(X)/length(X)
return(paste('The mean of',M))
}

mean.function(ny$Trip.Duration)   #Mean of New_York_City Bikeshares Trip Duration
mean.function(wash$Trip.Duration) #Mean of Washigton Bikeshares Trip Duration
mean.function(chi$Trip.Duration)  #Mean of Chicago Bikeshares Trip Duration
```

#NUMBER 1a ANSWER SUMMARY
#The mean of New York data trip duration is 899.68.
#The mean of Washington data trip duration is 1237.28.
#The mean ofChicago data trip duration is 936.24.

```{r USER TYPE THAT ACCUMULATED THE MOST TRIP DURATION.}
#1b User Tupe that accumulated the most Trip Duration
#Getting the User type that accumulated the most trip duration in each cities?

#New York
ny1=ny %>%            #Removing NA values for User Type in New York.
mutate_if(is.factor,trimws) %>% 
filter(User.Type!='') 

qplot(x=Trip.Duration, #Plotting User Type Trip duration for New York.
data=ny1,
binwidth=100,
color=I('blue')) +
facet_wrap(~User.Type)+
coord_cartesian(xlim =c (0,10000))+
xlab('Trip Duration For People in New York')+
ylab('Number of People')

by(ny1$Trip.Duration,ny1$User.Type,summary)
by(ny1$Trip.Duration,ny1$User.Type,sum)  #Getting New York Trip Duration count

#Washington
wash1=wash %>%        #Removing NA values for User Type in Washington
mutate_if(is.factor,trimws) %>% 
filter(User.Type!='') 

qplot(x=Trip.Duration, #Plotting User.Group Trip duration for Washington.
data=subset(wash1),
binwidth=100,
color=I('blue')) +
facet_wrap(~User.Type)+
coord_cartesian(xlim =c (0,15000))+
xlab('Trip Duration for People in  Washington') +
ylab('Number of People')

by(wash1$Trip.Duration,wash1$User.Type,summary)
by(wash1$Trip.Duration,wash1$User.Type,sum)   #Getting Washington Trip Duration count

#Chicago
chi1 =chi %>%     #Removing NA values for User Type in Chicago
mutate_if(is.factor,trimws) %>% 
filter(User.Type!='') 

qplot(x=Trip.Duration, #Plotting User.Group Trip duration for Chicago.
data=chi1,
binwidth=100,
color=I('blue')) +
facet_wrap(~User.Type)+
coord_cartesian(xlim =c (0,50000))+
xlab('Trip Duration for People in  Chicago') +
ylab('Number of People')

by(chi1$Trip.Duration,chi1$User.Type,summary)
by(chi1$Trip.Duration,chi1$User.Type,sum)   #Getting Chicago Trip Duration count

```

#NUMBER 1b ANSWERS SUMMARY
#New York
#Subcribers has a higher accumulated trip duration of 206864152 compared to Customer with an accumulated trip of 
#62021801 Just as the plot depicted.
#Washington
#Washington bike data,Customers have a higher accumlated trip duration of 208740308 compared to 
#subcribers accumulated trip duration of 162443678.
#Chicago
#Chicago bike data. Subcribers have a higher accumulated trip duration of 167054651 compared to customer and dependent
#having an accumulated trip duration of 113816825 and 311 respectively.

#NUMBER 2 START STATION AND END STATION.
#Plot and determine the start station and end station in each state that has the highest count.

```{r START STATION WITH THE HIGHEST COUNT}
#2a.Start Stations

#New York
qplot(x=Start.Station, #Plotting Start Station for New York.
data=subset(ny,!is.na(Start.Station)),
stat = ('count'), 
color=I('blue')) +
xlab('Start Station in New York')+
ylab("count")

ny %>%     #most popular start Station in New York.
count(Start.Station) %>% 
filter(n==max(n))  

#Washington
qplot(x=Start.Station, #Plotting Start Station for washington.
data=subset(wash,!is.na(Start.Station)),
stat = ('count'), 
color=I('blue')) +
xlab('Start Station in Washington')+
ylab("count")

wash %>%    #most popular start Station in Washington.
count(Start.Station) %>% 
filter(n==max(n))   

#Chicago
qplot(x=Start.Station, #Plotting Start Station for Chicago.
data=subset(chi,!is.na(Start.Station)),
stat = ('count'), 
color=I('blue')) +
xlab('Start Station in Chicago')+
ylab("count")

chi%>% 
count(Start.Station) %>%     #most popular start Station in Chicago.
filter(n==max(n))  

```

#NUMBER 2a ANSWER SUMMARY
#The most used start station in New York is Pershing Square North with a count of 3069 as shown in the plot.
#The most used start station in Washington is Columbus Circle/Union Station with a count of 5656 as shown in the plot. 
#The most used start station in Chicago is Streeter Dr and Grand Ave with a count of 6911 as shown in the plot.

```{r END STATION WITH THE HIGHEST COUNT}
#2b End Stations

#New York
qplot(x=End.Station, #Plotting Start Station for New York.
data=subset(ny,!is.na(End.Station)),
stat = ('count'), 
color=I('blue')) +
xlab('End Station in New York')+
ylab("count")

ny %>%     #most popular End Station in New York.
count(End.Station) %>% 
filter(n==max(n))  

#Washington
qplot(x=End.Station, #Plotting Start Station for washington.
data=subset(wash,!is.na(End.Station)),
stat = ('count'), 
color=I('blue')) +
xlab('End Station in Washington')+
ylab("count")

wash %>%    #most popular end Station in Washington.
count(End.Station) %>% 
filter(n==max(n))   

#Chicago
qplot(x=End.Station, #Plotting Start Station for washington.
data=subset(chi,!is.na(End.Station)),
stat = ('count'), 
color=I('blue')) +
xlab('End Station in Chicago')+
ylab("count")

chi %>%    #most popular end Station in Chicago.
count(End.Station) %>% 
filter(n==max(n)) 
```

#NUMBER 2b ANSWER SUMMARY
#The most used end station in New York is Pershing Square North with a count of 30777 as shown in the plot.
#The most used end station in Washington is Columbus Circle/Union Station with a count of 6048 as shown in the plot. 
#The most used end station in Chicago is Streeter Dr and Grand Ave with a count of 7512 as shown in the plot.

#NUMBER 3 START TIME AND END TIME
#Plot and determine the most month and day that occurs mostly in the start time and end time of the three cities?

```{r START MONTH AND DAY OF THE WEEK WITH HIGHEST COUNT}
#3a Start time month and Day of the week
#New York
#New York Start time Month
ny=ny %>%       #mutating the start time
mutate(Start.Time=ymd_hms(Start.Time))
head(ny)

ny=ny %>%       #mutating the start time and dividing by month
mutate(month.name=month(Start.Time, 
label = TRUE))

qplot(x=month.name,     #Plotting the months in the start time against counts
data=ny,
color=I('green'))+
xlab('New York months')+
ylab('counts')

ny %>%        #Getting the month with the highest count
count(month.name) %>% 
filter(n==max(n)) 

#New York Start time Day of the week
ny=ny%>%      #mutating the start time and dividing by day of the week
mutate(day_week=wday(Start.Time,
label = TRUE))

qplot(x=day_week,    #Plotting the days of the week in the start time against counts
data=ny,
color=I('green')) +
xlab('New York Days of the week')+
ylab('counts')
      
ny %>%        #Getting the day of the week with the highest count
count(day_week) %>% 
filter(n==max(n)) 

#Washington
#NWashington Start time Month
wash=wash %>%       #mutating the start time
mutate(Start.Time=ymd_hms(Start.Time),
End.Time=ymd_hms(End.Time))
head(wash)

wash=wash %>%       #mutating the start time and dividing by month
mutate(month.name=month(Start.Time, 
label = TRUE))

qplot(x=month.name,     #Plotting the months in the start time against counts
data=wash,
color=I('green'))+
xlab('Washington months')+
ylab('counts')

wash %>%        #Getting the month with the highest count
count(month.name) %>% 
filter(n==max(n)) 

#Washington Start time Day of the week
wash=wash%>%      #mutating the start time and dividing by day of the week
mutate(day_week=wday(Start.Time,
label = TRUE))

qplot(x=day_week,    #Plotting the days of the week in the start time against counts
data=wash,
color=I('green')) +
xlab('Washington Days of the week')+
ylab('counts')

wash %>%        #Getting the day of the week with the highest count
count(day_week) %>% 
filter(n==max(n))

#Chicago
#Chicago Start time Month
chi=chi %>%       #mutating the start time
mutate(Start.Time=ymd_hms(Start.Time),
End.Time=ymd_hms(End.Time))
head(chi)

chi=chi %>%       #mutating the start time and dividing by month
mutate(month.name=month(Start.Time, 
label = TRUE))

qplot(x=month.name,     #Plotting the months in the start time against counts
data=chi,
color=I('green'))+
xlab('Chicago months')+
ylab('counts')

chi %>%        #Getting the month with the highest count
count(month.name) %>% 
filter(n==max(n)) 

#Chicago Start time Day of the week
chi=chi%>%      #mutating the start time and dividing by day of the week
mutate(day_week=wday(Start.Time,
label = TRUE))

qplot(x=day_week,    #Plotting the days of the week in the start time against counts
data=chi,
color=I('green')) +
xlab('Chicago Days of the week')+
ylab('counts')

chi %>%        #Getting the day of the week with the highest count
count(day_week) %>% 
filter(n==max(n))
```

#NUMBER 3a ANSWER SUMMARY
# New York start time most populous month is June with a count of 76022.
# New York start time most populous day of the week is Wednesday with a count of 52087.
# Washington start time most populous month is June with a count of 68339.
# Washington start time most populous day of the week is Wednesday with a count of 48156.
# Chicago start time most populous month is June with a count of 98081.
# Chicago start time most populous day of the week is Tuesday with a count of 45912.

```{r END MONTH AND DAY OF THE WEEK WITH HIGHEST COUNT}
#3b End time month and Day of the week
#End Time
#New York End Time
ny=ny %>%       #mutating the end time
mutate(End.Time=ymd_hms(End.Time))
head(ny)

ny=ny %>%       #mutating the end time and dividing by month
mutate(month.name=month(End.Time, 
label = TRUE))

qplot(x=month.name,     #Plotting the months in the end time against counts
data=ny,
color=I('green'))+
xlab('New York months')+
ylab('counts')

ny %>%        #Getting the month with the highest count
count(month.name) %>% 
filter(n==max(n)) 

#New York Start time Day of the week
ny=ny%>%      #mutating the end time and dividing by day of the week
mutate(day_week=wday(End.Time,
label = TRUE))

qplot(x=day_week,    #Plotting the days of the week in the end time against counts
data=ny,
color=I('green')) +
xlab('New York Days of the week')+
ylab('counts')

ny %>%        #Getting the day of the week with the highest count
count(day_week) %>% 
filter(n==max(n)) 

#Washington
#NWashington End time Month
wash=wash %>%       #mutating the end time
mutate(Start.Time=ymd_hms(Start.Time),
End.Time=ymd_hms(End.Time))
head(wash)

wash=wash %>%       #mutating the end time and dividing by month
mutate(month.name=month(End.Time, 
label = TRUE))

qplot(x=month.name,     #Plotting the months in the end time against counts
data=wash,
color=I('green'))+
xlab('Washington months')+
ylab('counts')

wash %>%        #Getting the month with the highest count
count(month.name) %>% 
filter(n==max(n)) 

#Washington Start time Day of the week
wash=wash%>%      #mutating the end time and dividing by day of the week
mutate(day_week=wday(End.Time,
label = TRUE))

qplot(x=day_week,    #Plotting the days of the week in the end time against counts
data=wash,
color=I('green')) +
xlab('Washington Days of the week')+
ylab('counts')

wash %>%        #Getting the day of the week with the highest count
count(day_week) %>% 
filter(n==max(n))

#Chicago
#Chicago End time Month
chi=chi %>%       #mutating the end time
mutate(Start.Time=ymd_hms(Start.Time),
End.Time=ymd_hms(End.Time))
head(chi)

chi=chi %>%       #mutating the end time and dividing by month
mutate(month.name=month(End.Time, 
label = TRUE))

qplot(x=month.name,     #Plotting the months in the start time against counts
data=chi,
color=I('green'))+
xlab('Chicago months')+
ylab('counts')

chi %>%        #Getting the month with the highest count
count(month.name) %>% 
filter(n==max(n))

#Chicago End time Day of the week
chi=chi%>%      #mutating the end time and dividing by day of the week
mutate(day_week=wday(End.Time,
label = TRUE))

qplot(x=day_week,    #Plotting the days of the week in the end time against counts
data=chi,
color=I('green')) +
xlab('Chicago Days of the week')+
ylab('counts')

chi %>%        #Getting the day of the week with the highest count
count(day_week) %>% 
filter(n==max(n))
```

#NUMBER 3b ANSWER SUMMARY
# New York end time most populous month is June with a count of 76019.
# New York end time most populous day of the week is Wednesday with a count of 52070.
# Washington end time most populous month is June with a count of 68329.
# Washington end time most populous day of the week is Wednesday with a count of 48134.
# Chicago end time most populous month is June with a count of 98074.
# Chicago start time most populous day of the week is Tuesday with a count of 45887.


